---
title: "Activity 2 - Day 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading the packages 
```{r}

library(tidyverse)
library(tidymodels)
```


### Task3: Load the data 
```{r}
hfi <- readr::read_csv("https://www.openintro.org/data/csv/hfi.csv")
hfi

```

1. what are the dimensions of the data set what does each row represent? 
Ans. There are 1458 rows and 123 columns,it represents the Human Freedom Index 


## Filtering out the year 2016 from the dataset
```{r}

hfi_2016 <- hfi %>% filter(year == 2016) # using hfi_2016 as new object to store the values of the year 2016 from the original file
hfi_2016

```


## Linear Model for Prediction
```{r}
pf_model <- lm(pf_score ~ pf_expression_control, data = hfi_2016)
pf_model
```


```{r}

ggplot(hfi_2016, aes(x = pf_expression_control, y = pf_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)+
geom_point() + 
  labs(title = " Public Freedom - 2016 ",
              x = "Expression Control", y = " Scores" )

```


Q. Does the relationship looks linear? 
Ans. Yes, the linear predictor line looks strongly linear and would definitely be able to use this model to predict the personal freedom scores. 


### Task 4 

The smallest number that I ever got was 271.362, and my best one yet is 307.98. My neighbors have somewhat similar results. 


### Taks 5 

```{r}
 m1 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)
tidy(m1)
```
















