---
title: "Activity 2 - Day 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Loading the packages 
```{r}

library(tidyverse)
library(tidymodels)
```


### Task3: Load the data 
```{r}
hfi <- readr::read_csv("https://www.openintro.org/data/csv/hfi.csv")
hfi

```

1. what are the dimensions of the data set what does each row represent? 
Ans. There are 1458 rows and 123 columns,it represents the Human Freedom Index 


## Filtering out the year 2016 from the dataset
```{r}

hfi_2016 <- hfi %>% filter(year == 2016) # using hfi_2016 as new object to store the values of the year 2016 from the original file
hfi_2016

```


## Linear Model for Prediction
```{r}
pf_model <- lm(pf_score ~ pf_expression_control, data = hfi_2016)
pf_model
```


```{r}

ggplot(hfi_2016, aes(x = pf_expression_control, y = pf_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)+
geom_point() + 
  labs(title = " Public Freedom - 2016 ",
              x = "Expression Control", y = " Scores" )

```


Q. Does the relationship looks linear? 
Ans. Yes, the linear predictor line looks strongly linear and would definitely be able to use this model to predict the personal freedom scores. 


### Task 4 

The smallest number that I ever got was 271.362, and my best one yet is 307.98. My neighbors have somewhat similar results. 


### Taks 5 

```{r}
 m1 <- lm(pf_score ~ pf_expression_control, data = hfi_2016)
tidy(m1)
```


### Activity 2 - Day 2 

```{r}
a <- cor(hfi_2016$pf_expression_control, hfi_2016$pf_score)

```

The correlation coefficient between pf_expression_control and pf_Score is `r a`. This means that both are highly correlated. 


```{r}
glance(m1)
```

The Value of R-squared in the model is 0.714, which means that 71.4 percent of the variation can be explained by the model. 

```{r}
pf_m <- lm(hf_score ~ pf_expression_control, data = hfi_2016)
pf_m
```

```{r}
ggplot(hfi_2016, aes(x = pf_expression_control, y = hf_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)+
geom_point() + 
  labs(title = " Public Freedom - 2016 ",
              x = "Expression Control", y = " Hf - Scores" )
```

The slope is linear, which means both the values have a high correlation coefficient, therefore the linear equation for the model will be: hf_score = 5.0534 + 0.3684 * pf_expression_control. According to the equation, if the pf_expression_control is said to be 3, then if I go by the equation, the hf_score will be 6.15, which is a good score. 

Higher the pf_expression_control value, higher would be the the hf_score, I believe it is an overestimate because the residuals value are positive. 

```{r}
m1_aug <- augment(m1)
m1_aug
```


```{r}
ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```

The points appear to be random and not in a pattern in the residuals plot. The distributions are symmetric, however, there are outliers (2). 

```{r}
ggplot(data = m1_aug, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

Looking at the histogram, I believe that the model is a bit skewed towards the left, this also means the residuals are not linearly separated and most of the residuals are negative. 

